---
title: "어떤 연구가 좋을까?"
author: "Jae Kwan Koo"
output: slidy_presentation
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

## Contents

- TCGA in R
    - Install
    - Example - clinic data
    - Download data from GDC(Genomic Data Commons Data Portal)  

- 관심이 가는 주제  


## TCGA in R  

**TCGA(The Cancer Genome Atlas)**  

* The Cancer Genome Atlas (TCGA)는 미국 국립보건원(National Institutes of Health, NIH)에서 진행한 프로젝트  

* 여러 가지 암종에 대한 유전체/전사체/단백체 데이터를 수집 및 분석하여 암을 분자 수준에서 이해  

* 게놈 염기서열 분석과 생물정보학을 이용해 암의 원인이 되는 유전자 돌연변이를 분류하기 위해 2005년부터 시작된 프로젝트. 대부분의 유전체학 연구보다 많은 500개의 환자 샘플을 계획했고, 환자 샘플을 분석하기 위해 다른 기술을 사용했다.  

## Install  

```{r message=F,warning=FALSE}
# devtools::install_github("halpo/purrrogress")
# devtools::install_github("waldronlab/MultiAssayExperiment")
# devtools::install_github("PoisonAlien/maftools")

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

if (!requireNamespace("ComplexHeatmap", quietly = TRUE))
  BiocManager::install("ComplexHeatmap") # from bioconductor

if (!requireNamespace("TCGAbiolinks", quietly = TRUE))
  BiocManager::install("BioinformaticsFMRP/TCGAbiolinks") # from github

library(TCGAbiolinks)
library(MultiAssayExperiment)
library(maftools)
library(tidyverse)
library(ComplexHeatmap)
```  


## Example - clinic data  

```{r}
clinical <- GDCquery_clinic("TCGA-COAD")

DT::datatable(clinical, 
              options = list(pageLength = 5, autoWidth = TRUE))
```  

## 

```{r echo=FALSE, fig.width=18, fig.height=10, fig.align='center'}
null_count <- apply(clinical, 2, function(x) sum(is.na(x)))

data <- data.frame(col_names = null_count %>% names, count = null_count)
rownames(data)  <- NULL


data %>% ggplot(aes(x = col_names, y = count)) + 
  geom_bar(stat="identity") + 
  coord_flip() + 
  xlab("") + 
  ylab("") + 
  theme_bw()
```  

## Download data from GDC(Genomic Data Commons Data Portal)  

```{r eval=F}
# 1. Search data in GDC
query <- GDCquery(project="TCGA-LIHC",    # Liver and intrahepatic bile ducts(간, 간내 담관)
                  data.category="Transcriptome Profiling",
                  data.type="Gene Expression Quantification",
                  workflow.type="HTSeq - Counts")


# 2. Download from GDC repository
GDCdownload(query)


# 3. Make R object from the downloaded data
data <- GDCprepare(query)

# 4. Extract Gene expression matrix
library(SummarizedExperiment)

eset <- assay(data)

# 5. Save the matrix as .csv format
fwrite(x = eset, file = "LIHC.csv") # setwd() location
```  

##

```{r eval=T}
TCGAbiolinks:::getGDCprojects()$project_id    # possible project name

TCGAbiolinks:::getProjectSummary("TCGA-LIHC")
```  


## [관심이 가는 주제](https://ko.ovalengineering.com/quantitative-mri-radiomics-prediction-molecular-classifications-breast-cancer-subtypes-tcgatcia-data-870194#menu-8)  

* 유방암은 북미 여성에서 가장 흔히 진단되는 암으로 여성의 암 사망 원인 중 두 번째로 많이 발생  

* 수용체 상태에 따라 유방암은 여러 가지 아형으로 분류 될 수 있음.  

<br>
<br>


[TCGA Cancers Selected for Study](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga/studied-cancers)  

What have TCGA researchers learned about breast cancer?  

* The cancer can be categorized into four molecular subtypes: `HER2-enriched`, `Luminal A`, `Luminal B`, and `Basal-like`.
Each subtype is associated with a unique panel of mutated genes.  

* Basal-like(기저형 아형) subtype shares many genetic features with high-grade serous ovarian cancer(난소 암), suggesting that the cancers have a common molecular origin and may share therapeutic opportunities, such as:  
    * A drug that inhibits blood vessel growth, cutting off the blood supply to the tumor.  

    * Bioreductive drugs, which are inactive drugs that become toxic to cancer cells under low oxygen conditions.



## Refer  

* TCGA  

    * https://bioconductor.org/packages/release/bioc/html/TCGAbiolinks.html  
    
    * [National cancer institute GDC Data Portal](https://portal.gdc.cancer.gov/projects)  

    * https://m.blog.naver.com/PostView.nhn?blogId=cjh226&logNo=221320056674&proxyReferer=https:%2F%2Fwww.google.com%2F  

    * https://rpubs.com/tiagochst/TCGAworkshop  
    
    * [TCGA Barcode](http://blog.naver.com/PostView.nhn?blogId=cjh226&logNo=220992617831&categoryNo=18&parentCategoryNo=0&viewDate=&currentPage=1&postListTopCurrentPage=1&from=postView)  
    
    * [TCGA의 간단한 설명을 한 블로그](https://medium.com/biosupermarket/%EB%8D%B0%EC%9D%B4%ED%84%B0-%EB%82%98%EB%93%A4%EC%9D%B4-tcga-target-gtex-1-tcga-target-gtex-32b3a78de709)  
    
    * [GDC query](https://rdrr.io/bioc/TCGAbiolinks/man/GDCquery.html)  
    












